# ANOVA analysis to identify metabolites/lipids that have significant cell type:age interaction in DESI-MSI data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list=ls())
library(tidyverse)
library(data.table)
library(RColorBrewer)
library(mixOmics)
library(ComplexHeatmap)
source('data_loader.R')
source('deconvolution_functions.R')
```

# Data loading and processing

Load data and write functions to convert it into a matrix for follow-up analyses:

```{r}
#load Effect Size table
ssDT <- fread('../../Output_Data//20210512_DESI_decomposition_single_sample.csv')
```

```{r}
dt2matrix <- function(inputDT){
  inM <- as.matrix(inputDT[,-1])
  row.names(inM) <- inputDT$sampleID
  return(inM)
}

allcCastMat <- dt2matrix(
  dcast.data.table(ssDT[cell!='other'],
                   sampleID~cell+peak,
                   value.var = 'mean'))

cellCastMat <- dt2matrix(
  dcast.data.table(ssDT[cell!='other'],
                   paste0(sampleID,"_",cell)~peak,
                   value.var = 'mean'))

gfapCastMat <- dt2matrix(
  dcast.data.table(ssDT[cell=='gfap'],
                   sampleID~cell+peak,
                   value.var = 'mean'))
```


# Analysis of interactions between age and cell type using ANOVA

Previous efforts to use unsupervised learning sucked. The issue (classic) is that
we have way too many features and not enough samples. Perform ANOVA using a simple 
linear model (try later with GAMs), to perform feature selection. Cell type and age 
are encoded as factors. Response is the mean level per sample.

```{r}
ssDTanovaCellPeak <- rbindlist(
  lapply(
    ssDT[,.N,peak]$peak,
    function(peakID){
      as.data.table(
        anova(
          lm(mean~class*cell -1,
             data=ssDT[peak==peakID])
        )[5],keep.rownames = T)[,peak:=peakID]
    }
  ))[order(`Pr(>F)`)][!is.na(`Pr(>F)`)]
```



```{r}
ssDTanovaCellPeakInt <- copy(ssDTanovaCellPeak[rn=='class:cell'])
ssDTanovaCellPeakInt[,fdr:=p.adjust(`Pr(>F)`,method = 'fdr')]
ssDTanovaCellPeakInt[fdr < .2]
```

```{r}
library(ggplot2)
library(ggthemes)
library(ggbeeswarm)

ssDT[,age:="Young"]
ssDT[class==1,age:="Old"]

ssDT[,cellS:='o']
ssDT[cell=='both',cellS:='aNSC']
ssDT[cell=='gfap',cellS:='qNSC']
ssDT[cell=='ki67',cellS:='NPC']

ssDT[,scaled_y:=scale(mean),peak]

ssDT$age <- factor(ssDT$age, levels = c("Young", "Old"))
ssDT$cellS <- factor(ssDT$cellS, levels = c("qNSC", "aNSC", "NPC"))

ggplot(ssDT[cell!='other'][
  peak %in% ssDTanovaCellPeakInt[fdr < .2]$peak
],
aes(x=cellS,
    color=age,
    y=scaled_y)) + 
  geom_boxplot(outlier.shape = NA) + facet_grid(.~peak) + 
  geom_point(position = position_jitterdodge(), alpha = 0.7,
             aes(shape = cellS), size = 2.5)+
  theme_classic()+
  theme(legend.position = 'bottom') + 
  labs(fill="") + xlab('cell type') +
  ylab("normalized metabolite intensity") +
  ggtitle("Lipid peaks with significant age:cell interactions") +
  scale_color_manual(values = c("darkgoldenrod", "maroon"))+
  scale_shape_manual(values=c(17,16,18))+
  theme(text=element_text(size = 12, face = "plain"),
        axis.title = element_text(colour = "black"))
ggsave(filename = paste0("../../Figure_Panels/Fig_S3f.pdf"),
       width = 6, height = 5, useDingbats=FALSE)
```


