---
title: "R Notebook"
---
## _In silico_ admixture deconvolution to validate linear regression approached used in DESI lipidomics data
## 4 cell type-specific lipidomics profile
### Activated NSC - Our study
### Quiescent NSC - Our study
### Cultured neurons - Fitzner et al. 2020 https://www.sciencedirect.com/science/article/pii/S2211124720311219#app2
### Cultured oligodendrocytes - Fitzner et al. 2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
rm(list=ls())
library(ragg)
library(tidyverse)
library(data.table)
source('data_loader.R')
source('deconvolution_functions.R')
```

qNSC and aNSC data
```{r}
load("../../Input_Data/LC_Exp1_for_admixture.Rdata")
```

Neuron and oligodendrocytes data
```{r}
load("../../Input_Data/Neuron_Oligo_for_admixture.Rdata")
```

Color scheme
```{r}
col.ls <- c("#7D65A5", "#81B486", "darkorange4", "deepskyblue4")
```


Combine published and internal datasets for testing accuracy of cell-mixture
deconvolution

```{r}
invitroDT <- melt.data.table(as.data.table(LC.AQ),
                             id.vars = 'ID_string')
invitroDT[,status:='Quiescent']
invitroDT[grep("Activ",variable),status:='Activated']
invitroDT[,subj:=gsub("Activated|Quiescent","",variable)]
```


```{r}
NEUoligoDT <- melt.data.table(as.data.table(BR.Neu.Oligo),
                              id.vars = 'ID_string')
NEUoligoDT[,status:='Neuron']
NEUoligoDT[grep("Oligo",variable),status:='Oligodendrocytes']
NEUoligoDT[,subj:=gsub("Oligodendrocytes|Neurons","",variable)]
```


```{r}
vectorNospace <- function(inVec){
  sapply(inVec,function(i)
    gsub(" +","",i))
}

listSortCollapse <- function(inlist){
  paste(sort(vectorNospace(inlist)),collapse = ',')
}

sortCollapse <- function(instring){
  listSortCollapse(strsplit(instring,",")[[1]])
}


NEUoligoDT[,newID:=sapply(NEUoligoDT$ID_string,
                          function(i)
                            listSortCollapse(i))]


invitroDT[,newID:=sapply(invitroDT$ID_string,
                         function(i)
                           listSortCollapse(i))]


combDT <- rbind(
  NEUoligoDT,invitroDT)[,mean(value,na.rm=T),
                        .(variable,newID)][!is.na(V1)]

```

## Create synthetic admixtures using both in silico cells and real in vitro data

Combine cell peak profiles using arbitrary proportions into admixtures


```{r}
#for every subject get all cells, then generate n admixtures
mixEngine <- function(inputCastDT){
  
  #generate random proportion
  propV <- sample(1:100,ncol(inputCastDT)-1,replace = T)
  propV <- propV/sum(propV)
  
  #compute mix
  return(data.table(
    rn=inputCastDT$rn,
    mix=as.matrix(inputCastDT[,-1]) %*% propV,
    Activated = propV[1],
    Neurons = propV[2],
    Oligodendrocytes = propV[3],
    Quiescent = propV[4]#this is hardcoded now,fix later
  ))
}

mixSbjRunner <- function(inputDT,
                         subject,
                         mixNumber){
  
  #cast data.table
  castDT <- dcast.data.table(inputDT[subj==subject],rn~status)
  
  #generate proportion vector
  return(
    rbindlist(lapply(1:mixNumber,function(i)
      mixEngine(castDT)[,n:=i])
    )[,sbj:=subject])
}

mixSbjWrapper <- function(inputDT,...){
  
  #for each subject compute admixtures
  return(
    rbindlist(
      lapply(inputDT[,.N,subj]$subj,
             function(sID){
               mixSbjRunner(inputDT,sID,...)
             })))
}
```


Select peaks that we care about [one with representation across all]

```{r}
admixture_input_DT <- copy(
  combDT[newID %in% combDT[,.N,newID][order(N)][N==42]$newID]
)

admixture_input_DT[,Lsubj:=gsub(
  "Activated|Quiescent|Neurons|Oligodendrocytes",
  "",
  variable)]
```

Because the set have mismatched samples, we are going to match them by number
and assign one randomly paired sample of neurons and oligo to each mouse. Setting
seed to 3333 to avoid issues with reproducibility in this step.

```{r}
set.seed(3333)

invitroSamples <- sample(admixture_input_DT[grep("Activ",variable),
                                            .N,Lsubj]$Lsubj,9)

neuronSamples  <- sample(admixture_input_DT[grep("Neuron",variable),
                                            .N,Lsubj]$Lsubj,9)

oligosSamples  <- sample(admixture_input_DT[grep("Oligo",variable),
                                            .N,Lsubj]$Lsubj,9)

match_id_DT <- as.data.table(rbind(
  cbind(invitroSamples,invitroSamples),
  cbind(neuronSamples,invitroSamples),
  cbind(oligosSamples,invitroSamples))
)
setnames(match_id_DT,c("Lsubj","subj"))

```


```{r}
admixture_input_DT <- admixture_input_DT[match_id_DT,o=.(Lsubj)]
```

```{r}
admixture_input_DT[,status:='Quiescent']
admixture_input_DT[grep("Activ",variable),status:='Activated']
admixture_input_DT[grep("Activ",variable),status:='Activated']
admixture_input_DT[grep("Neur",variable),status:='Neurons']
admixture_input_DT[grep("Oligo",variable),status:='Oligodendrocytes']
```

Datasets are on different scales, let's turn them into Z-scores

```{r}
admixture_input_DT[,value:=scale(V1),.(newID,status)]
admixture_input_DT[,rn:=newID]
```

Combine data into admixtures [set random seed for reproducibility]

```{r}
set.seed(3333)
invitro_neurons_MixDT <- mixSbjWrapper(admixture_input_DT,mixNumber=10)
```

```{r}
mixPropDT <- unique(
  invitro_neurons_MixDT[,c('Activated',
                           'Neurons',
                           'Oligodendrocytes',
                           'Quiescent',
                           'n','sbj')])
mixPropDT[,sample:="AAA"]

for(i in 1:nrow(mixPropDT)){
  mixPropDT[i,sample:=paste(sbj,n,sep = "_")]
}

mixPropDT <- mixPropDT[,c('sample',
                          'Activated',
                          'Neurons',
                          'Oligodendrocytes',
                          'Quiescent')]
```

```{r}
mixPeakDT <- unique(invitro_neurons_MixDT[,c('rn','mix.V1','n','sbj')])
mixPeakDT[,sample:="AAA"]

for(i in 1:nrow(mixPeakDT)){
  mixPeakDT[i,sample:=paste(sbj,n,sep = "_")]
}
mixPeakDT <- dcast.data.table(mixPeakDT,sample~rn,value.var = 'mix.V1')
```

We now have a table of peak values from a mixed tissue and the corresponding 
proportions that are behind the values in the mix


## Test decomposition methods on syntethic mixtures

run csSAM and NNLS on the admixtures as groups

```{r}
csSAMfitYadMIX <- csSAMcsFit(mixPeakDT[grep("Y",sample)],
                             mixPropDT[grep("Y",sample)])

csSAMfitOadMIX <- csSAMcsFit(mixPeakDT[grep("O",sample)],
                             mixPropDT[grep("O",sample)])
```


then run it for a single subject [functionalize this for further comparisons]

```{r}
#functionalized single sample deconvolution runner
singleSampleDecRunner <- function(inputPeakDT,
                                  inputPropDT){
  
  subjV <- unique(
    sapply(inputPeakDT[,.N,sample]$sample,
           function(i)
             paste(strsplit(i,"_")[[1]][1:3],collapse =  "_")))
  
  subjectOutDT <- lapply(
    subjV,
    function(sampleID){
      
      #run both methods
      csSAMfitTemp <- csSAMcsFit(
        inputPeakDT[grep(sampleID,sample)],
        inputPropDT[grep(sampleID,sample)])
      
      #label them
      csSAMfitTemp[,method:='csSAM']
      
      return(
        csSAMfitTemp[,sampleID:=sampleID]
      )
    }
  )
  
  return(
    rbindlist(
      subjectOutDT
    )
  )
}
```


```{r}
adMixsingleSubjectDecompositionDT <- singleSampleDecRunner(mixPeakDT,mixPropDT)
```


merge decomposition results with the original values

```{r}
cmpDT <- admixture_input_DT[adMixsingleSubjectDecompositionDT,
                            o=.(status==cell,
                                subj==sampleID,
                                rn==peak)]
```


First plot: truth
```{r}
library(ggpubr)
library(ggthemes)

cmpDT[,Age:='Old']
cmpDT[grep("Y",subj),Age:='Young']

cmpDT$Age <- factor(cmpDT$Age, levels = c("Young", "Old"))

cmpDT$status <- factor(cmpDT$status, levels = c("Activated", "Quiescent", "Neurons", "Oligodendrocytes"))

ggplot(cmpDT[method=='csSAM'],
       aes(x=value,
           y=mean
       )) + 
  geom_point(aes(color=status), alpha = 0.3) + 
  facet_grid( ~ status)+
  scale_color_manual(values = col.ls)+
  theme(legend.position = 'bottom') + 
  xlab('Measured lipid intensity') +
  ylab("Estimated lipid intensity")+
  theme_classic() +
  theme(axis.text = element_text(colour = "black", 
                                 face = "plain",
                                 size = 8))+
  stat_cor(label.sep = "\n", size = 2.5)+
  theme(legend.position = "none")
ggsave("../../Figure_Panels/EDFig.6a.pdf", width = 6, height = 2, useDingbats=FALSE)
```

Perform a negative control experiment: permute proportions [n times] and assess
results

```{r}
permute_prop_DT <- function(inputDT){
  
  nCells   <- dim(inputDT)[1]*(dim(inputDT)[2]-1)
  inputMat <- t(
    apply(
      matrix(
        sample(nCells),
        ncol=4
      ),1,
      function(i) i/sum(i)
    ))
  row.names(inputMat) <- inputDT$sample
  
  outDT  <- as.data.table(
    inputMat,
    keep.rownames = T)
  setnames(outDT,colnames(inputDT))
  
  return(outDT)
}
```




```{r}
set.seed(3333)

adMixPermutationsDT <- rbindlist(
  lapply(1:100,
         function(nP)
           singleSampleDecRunner(
             inputPeakDT = mixPeakDT,
             inputPropDT = permute_prop_DT(mixPropDT))[,nP:=nP]))
```


merge decomposition results with the original values

```{r}
permuationsCmpDT <- admixture_input_DT[adMixPermutationsDT,
                                       o=.(status==cell,
                                           subj==sampleID,
                                           rn==peak)]
permuationsCmpDT[,Age:='Old']
permuationsCmpDT[grep("Y",subj),Age:='Young']

permuationsCmpDT$Age <- factor(permuationsCmpDT$Age,
                               levels = c("Young", "Old"))
```


Second plot: comparison to permutations


```{r}
permuationsCmpDT$status <- factor(
  permuationsCmpDT$status, levels = c("Activated", 
                                      "Quiescent", 
                                      "Neurons", 
                                      "Oligodendrocytes"))

cmpDT$status <- factor(cmpDT$status, levels = c("Activated", 
                                                "Quiescent", 
                                                "Neurons", 
                                                "Oligodendrocytes"))
ggplot(permuationsCmpDT[,cor(value,mean),
                        .(method,status,nP)][method=='csSAM'],
       aes(x=V1,y=..count.., fill=status)) + 
  
  geom_density(alpha = 0.75) + 
  
  scale_fill_manual(values = col.ls) + 
  
  xlab("Correlation coefficient between estimated and measured lipid intensity") +
  labs(fill='cell type')+
  
  theme(legend.position = 'none') +
  facet_grid(.~status) +
  
  geom_vline(data = cmpDT[method=='csSAM'][
    ,cor(value,mean),
    .(method,status)],
    aes(xintercept=V1,color=status),linetype=2) +
  scale_color_manual(values = col.ls) + 
  
  labs(fill='cell type')+
  labs(color='cell type') +
  
  theme_classic() + 
  theme(axis.text = element_text(colour = "black", 
                                 face = "plain",
                                 size = 8))+
  theme(legend.position = 'none')
ggsave("../../Figure_Panels/EDFig.6b.pdf", width = 6, height = 2, useDingbats=FALSE)
```






