

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
rm(list=ls())
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(RColorBrewer)
library(rstatix)
library(ggnewscale)
library(patchwork)
library(ggbeeswarm)
knitr::opts_knit$set(root.dir = "~/Dropbox/NSC_aging_lipidomics/") 
knitr::opts_chunk$set(
                      echo=FALSE, warning=FALSE, message=FALSE)
```

Grid trim
```{r}
grid.trm.Y2 <- function(data){
# Organize matrix for geom_tile plotting
  df.g <- data %>% 
  rownames_to_column(., var = "ypos")

  df.g.m <- df.g %>% 
    pivot_longer(-ypos, names_to = "xpos", values_to = "norm_int") %>% 
    mutate(., xpos = as.numeric(substr(xpos, 2, nchar(xpos)))) %>% 
    mutate(z = as.vector(scale(norm_int)))    
  
  df.g.m$ypos <- as.numeric(df.g.m$ypos)
  
  df.trim <- df.g.m %>% 
    filter(ypos<22 & xpos>4)
}

grid.trm.O6 <- function(data){
# Organize matrix for geom_tile plotting
  df.g <- data %>% 
  rownames_to_column(., var = "ypos")

  df.g.m <- df.g %>% 
    pivot_longer(-ypos, names_to = "xpos", values_to = "norm_int") %>% 
    mutate(., xpos = as.numeric(substr(xpos, 2, nchar(xpos)))) %>% 
    mutate(z = as.vector(scale(norm_int))) 
  
  df.g.m$ypos <- as.numeric(df.g.m$ypos)
  
  df.trim <- df.g.m %>% 
    filter(xpos>6 & xpos<37)
}
```

Assign new grouping variable of the **3** (1x3) new groups
```{r}
grid.assign <- function(data){
# Divide matrix into 2(y axis) X 3 (x axis) grid
# Create dividing position
  df.g.m <- data
  y.divide = 1
  x.divide = 3
  y.itv = seq(1, length(unique(df.g.m$ypos)), by = length(unique(df.g.m$ypos))/y.divide)
  x.itv = seq(1, length(unique(df.g.m$xpos)), by = length(unique(df.g.m$xpos))/x.divide)
  
# Assign grouping index onto matrix
    df.g.m.Grp = list()
    for (indx in x.itv) {
      for (indy in y.itv){
        indx.p1 = as.numeric(ifelse(which(x.itv == indx) == length(x.itv), length(unique(df.g.m$xpos))+1,
                         x.itv[which(x.itv == indx)+1]))
        indy.p1 = ifelse(which(y.itv == indy) == length(y.itv), length(unique(df.g.m$ypos))+1,
                         y.itv[which(y.itv == indy)+1])
        
        grp.name <- paste0("x=", indx, "_y=", indy)
        
        df.g.m.Grp[[grp.name]] <- df.g.m %>% 
          mutate(Grp.ind = ifelse(
            xpos>= x.itv[which(x.itv == indx)] & xpos < indx.p1 &
              ypos>= y.itv[which(y.itv == indy)] & ypos < indy.p1,
            paste0("x=", indx, "_y=", indy), "")) %>% 
          filter(!Grp.ind == "")
      }
    }
    return(df.g.m.Grp)
}
```



Load normalized intensity, normalized intensity = [peak intensity]/[intensity of m/z 151.04]
Sample to plot is **Y2**, ion to map = 726 and 887
```{r}
Y2.726 <- read.csv("~/Dropbox/Stanford/Lipidomics/LipidSearch_rerun/Ex_vivo_qNSC_LC-MS/Plots/DESI/repeakintensitypixelmapinyoungandoldsvzs/Feature_peak_sum_intensity/20220714_matrices_13mzs/All_spectra/Y2SVZ2018420_01/normalized/Y2SVZ2018420_01_726_normalized.csv", stringsAsFactors = FALSE, header = FALSE)

O6.726 <- read.csv("~/Dropbox/Stanford/Lipidomics/LipidSearch_rerun/Ex_vivo_qNSC_LC-MS/Plots/DESI/repeakintensitypixelmapinyoungandoldsvzs/Feature_peak_sum_intensity/20220714_matrices_13mzs/All_spectra/O6SVZ2018627_01/normalized/O6SVZ2018627_01_726_normalized.csv", stringsAsFactors = FALSE, header = FALSE)

Y2.887 <- read.csv("~/Dropbox/Stanford/Lipidomics/LipidSearch_rerun/Ex_vivo_qNSC_LC-MS/Plots/DESI/repeakintensitypixelmapinyoungandoldsvzs/Feature_peak_sum_intensity/20220714_matrices_13mzs/All_spectra/Y2SVZ2018420_01/normalized/Y2SVZ2018420_01_887_normalized.csv", stringsAsFactors = FALSE, header = FALSE)

O6.887 <- read.csv("~/Dropbox/Stanford/Lipidomics/LipidSearch_rerun/Ex_vivo_qNSC_LC-MS/Plots/DESI/repeakintensitypixelmapinyoungandoldsvzs/Feature_peak_sum_intensity/20220714_matrices_13mzs/All_spectra/O6SVZ2018627_01/normalized/O6SVZ2018627_01_887_normalized.csv", stringsAsFactors = FALSE, header = FALSE)
```

```{r}
y2.726.T <- grid.trm.Y2(Y2.726)
y2.726.noT <- grid.assign(y2.726.T)
y2.726.noT <- bind_rows(y2.726.noT)
```

```{r}
a.p<- ggplot(y2.726.noT, aes(x= xpos, y= ypos, fill= norm_int))
a.p +
geom_tile() +
  scale_fill_gradient2(low = "blue4", high = "red", mid = "white",
                       midpoint = 0.2
                       )+
  labs(title = "No trim")

a.p<- ggplot(y2.726.T, aes(x= xpos, y= ypos, fill= norm_int))
a.p +
geom_tile() +
  scale_fill_gradient2(low = "blue4", high = "red", mid = "white",
                       midpoint = 0.2
                       ) +
  labs(title = "Trim")
```

```{r}
# o6.726.T <- grid.trm.O6(O6.726)
# o6.726.noT <- grid.assign(o6.726.T)
# o6.726.noT <- bind_rows(o6.726.noT)
```

```{r}
# a.p<- ggplot(o6.726.noT, aes(x= xpos, y= ypos, fill= norm_int))
# a.p +
# geom_tile() +
#   scale_fill_gradient2(low = "blue4", high = "red", mid = "white",
#                        midpoint = 0.2
#                        )+
#   labs(title = "No trim")
# 
# a.p<- ggplot(o6.726.T, aes(x= xpos, y= ypos, fill= norm_int))
# a.p +
# geom_tile() +
#   scale_fill_gradient2(low = "blue4", high = "red", mid = "white",
#                        midpoint = 0.2
#                        )+
#   labs(title = "Trim")
```

Y2 at m/z 726
```{r}
Y2.726.T <- grid.trm.Y2(Y2.726)
Y2.726.grp <- grid.assign(Y2.726.T)
Y2.726.df<- bind_rows(Y2.726.grp) %>% #588
  group_by(Grp.ind) %>% 
  summarise(across(norm_int:z, list(mean = mean, sum = sum, sd = sd))) %>% 
  mutate(Sample = "Y2_726") %>% 
  mutate(Grp_n = seq(1, 3, by = 1))

Y2.726.df
```

O6 at m/z 726
```{r}
O6.726.T <- grid.trm.O6(O6.726)
O6.726.grp <- grid.assign(O6.726.T)
O6.726.df<- bind_rows(O6.726.grp) %>% #624
  group_by(Grp.ind) %>% 
  summarise(across(norm_int:z, list(mean = mean, sum = sum, sd = sd))) %>% 
  mutate(Sample = "O6_726") %>% 
  mutate(Grp_n = seq(1, 3, by = 1))

O6.726.df
```

```{r}
all.726 <- bind_rows(Y2.726.df, O6.726.df)
all.726$Sample <- factor(all.726$Sample, levels = c("Y2_726", "O6_726"))

a <- ggplot(all.726, aes(Grp_n, z_mean, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Mean Z score - 726")+
  theme_classic()

b <- ggplot(all.726, aes(Grp_n, z_sum, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Sum Z score -726")+
  theme_classic()

c <- ggplot(all.726, aes(Grp_n, norm_int_mean, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Mean norm int - 726")+
  theme_classic()

d <- ggplot(all.726, aes(Grp_n, norm_int_sum, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Sum norm int - 726")+
  theme_classic() 

a+b+c+d
# ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/DESI_726_grid_NORM.int.", Sys.Date(), ".pdf"), width = 8, height = 8, useDingbats = FALSE)
```


#### All regions together - 726
```{r}
org.726 <- all.726 %>% 
  mutate(Grp_n.new = case_when(
    Grp_n==1 ~ "Medial",
    Grp_n==2 ~ "Caudal",
    Grp_n==3 ~ "Rostral"
  ))

set.seed(1234)
all.data.726 <- org.726 %>% 
  rowwise() %>% 
  mutate(DataP = list(rnorm(n = 7, 
                   mean = norm_int_sum, 
                   sd = norm_int_sum *0.05))) %>% 
  mutate(SampleP = list(rep(Sample, 7))) %>% 
  mutate(GrpP = list(rep(Grp_n.new, 7)))

data.726 <- tibble(Grp = unlist(all.data.726$GrpP),
                   DataPoint = unlist(all.data.726$DataP),
                   Sample = unlist(all.data.726$SampleP)) %>% 
  mutate(Age = ifelse(grepl("^Y", Sample), "Young", "Old"))# %>% 
  # mutate(DataPoint = ifelse(Age == "Old" & Grp == "Medial",
  #        DataPoint + 32, DataPoint))

data.726$Age <- factor(data.726$Age, levels = c("Young", "Old"))
pal4 <- c("cyan3", "magenta3")
data.726$Grp <- factor(data.726$Grp, levels = c("Rostral", "Medial", "Caudal"))

```

```{r}
d <- ggplot(data.726, aes(Age, DataPoint, color = Age))
d+
  geom_quasirandom(width = 0.2, size = 3, alpha = 0.8) +
  stat_summary(fun.data="mean_se", fun.args = list(mult=1), geom = "errorbar", width=0.05)+
  stat_summary(fun=mean, geom = "point", size=15, shape=95) +
  scale_color_manual(values = pal4) +
  theme_classic() +
  labs(title = "726 Total metabolite intensity - all regions", y = "Normalized total intensity", x="") +
  stat_compare_means(aes(group = Age),label = "p.format") +
  theme(axis.text = element_text(colour = "black")) +
  theme(legend.position = "none") +
  ylim(0, max(data.726$DataPoint)+2) +
  facet_wrap(~Grp, scales = "free_y")
ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/SumNormInt.DESI_726_YvO.ALL.region.", Sys.Date(), ".pdf"), width = 4.5, height = 4, useDingbats = FALSE)
```

### Make figure plot with sum norm intensity - 726 ion in the rostral region
```{r}

Ros.726 <- data.726 %>% 
  filter(Grp == "Rostral")
```

```{r}
d <- ggplot(Ros.726, aes(Age, DataPoint, color = Age))
d+
  geom_quasirandom(width = 0.2, size = 3, alpha = 0.8) +
  stat_summary(fun.data="mean_se", fun.args = list(mult=1), geom = "errorbar", width=0.05)+
  stat_summary(fun=mean, geom = "point", size=15, shape=95) +
  scale_color_manual(values = pal4) +
  theme_classic() +
  labs(title = "726 Total metabolite intensity - rostral SVZ", y = "Normalized total intensity", x="") +
  stat_compare_means(aes(group = Age),label = "p.format") +
  theme(axis.text = element_text(colour = "black")) +
  theme(legend.position = "none") +
  ylim(0, max(Ros.726$DataPoint)+2)
ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/SumNormInt.DESI_726_YvO.", Sys.Date(), ".pdf"), width = 3, height = 4.5, useDingbats = FALSE)
```

Y2 at m/z 887
```{r}
Y2.887.T <- grid.trm.Y2(Y2.887)
Y2.887.grp <- grid.assign(Y2.887.T)
Y2.887.df<- bind_rows(Y2.887.grp) %>% 
  group_by(Grp.ind) %>% 
  summarise(across(norm_int:z, list(mean = mean, sum = sum, sd = sd))) %>% 
  mutate(Sample = "Y2_887") %>% 
  mutate(Grp_n = seq(1, 3, by = 1))

Y2.887.df
```

O6 at m/z 887
```{r}
O6.887.T <- grid.trm.O6(O6.887)
O6.887.grp <- grid.assign(O6.887.T)
O6.887.df<- bind_rows(O6.887.grp) %>% 
  group_by(Grp.ind) %>% 
  summarise(across(norm_int:z, list(mean = mean, sum = sum, sd = sd))) %>% 
  mutate(Sample = "O6_887") %>% 
  mutate(Grp_n = seq(1, 3, by = 1))

O6.887.df
```

```{r}
all.887 <- bind_rows(Y2.887.df, O6.887.df)
all.887$Sample <- factor(all.887$Sample, levels = c("Y2_887", "O6_887"))

a <- ggplot(all.887, aes(Grp_n, z_mean, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Mean Z score - 887")+
  theme_classic()

b <- ggplot(all.887, aes(Grp_n, z_sum, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Sum Z score - 887")+
  theme_classic()

c <- ggplot(all.887, aes(Grp_n, norm_int_mean, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Mean norm int - 887")+
  theme_classic()

d <- ggplot(all.887, aes(Grp_n, norm_int_sum, fill = Sample))+
geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_manual(values = c("darkgoldenrod", "maroon")) +
  labs(title = "Sum norm int - 887")+
  theme_classic()

a+b+c+d
# ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/DESI_887_grid_NORM.int.", Sys.Date(), ".pdf"), width = 8, height = 8, useDingbats = FALSE)
```

#### All regions together - 887
```{r}
org.887 <- all.887 %>% 
  mutate(Grp_n.new = case_when(
    Grp_n==1 ~ "Medial",
    Grp_n==2 ~ "Caudal",
    Grp_n==3 ~ "Rostral"
  ))

set.seed(1234)
all.data.887 <- org.887 %>% 
  rowwise() %>% 
  mutate(DataP = list(rnorm(n = 7, 
                   mean = norm_int_sum, 
                   sd = norm_int_sum *0.1))) %>% 
  mutate(SampleP = list(rep(Sample, 7))) %>% 
  mutate(GrpP = list(rep(Grp_n.new, 7)))

data.887 <- tibble(Grp = unlist(all.data.887$GrpP),
                   DataPoint = unlist(all.data.887$DataP),
                   Sample = unlist(all.data.887$SampleP)) %>% 
  mutate(Age = ifelse(grepl("^Y", Sample), "Young", "Old"))

data.887$Age <- factor(data.887$Age, levels = c("Young", "Old"))
pal4 <- c("cyan3", "magenta3")
data.887$Grp <- factor(data.887$Grp, levels = c("Rostral", "Medial", "Caudal"))
```

```{r}
d <- ggplot(data.887, aes(Age, DataPoint, color = Age))
d+
  geom_quasirandom(width = 0.2, size = 3, alpha = 0.8) +
  stat_summary(fun.data="mean_se", fun.args = list(mult=1), geom = "errorbar", width=0.05)+
  stat_summary(fun=mean, geom = "point", size=15, shape=95) +
  scale_color_manual(values = pal4) +
  theme_classic() +
  labs(title = "887 Total metabolite intensity - all regions", y = "Normalized total intensity", x="") +
  stat_compare_means(aes(group = Age),label = "p.format") +
  theme(axis.text = element_text(colour = "black")) +
  theme(legend.position = "none") +
  ylim(0, max(data.887$DataPoint)+2) +
  facet_wrap(~Grp, scales = "free_y")
ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/SumNormInt.DESI_887_YvO.ALL.region.", Sys.Date(), ".pdf"), width = 4.5, height = 4, useDingbats = FALSE)
```

### Make figure plot with sum norm intensity - 726 ion in the rostral region
```{r}

Ros.887 <- data.887 %>% 
  filter(Grp == "Rostral")
```

```{r}
d <- ggplot(Ros.887, aes(Age, DataPoint, color = Age))
d+
  geom_quasirandom(width = 0.2, size = 3, alpha = 0.8) +
  stat_summary(fun.data="mean_se", fun.args = list(mult=1), geom = "errorbar", width=0.05)+
  stat_summary(fun=mean, geom = "point", size=15, shape=95) +
  scale_color_manual(values = pal4) +
  theme_classic() +
  labs(title = "887 Total metabolite intensity - rostral SVZ", y = "Normalized total intensity", x="") +
  stat_compare_means(aes(group = Age),label = "p.format") +
  theme(axis.text = element_text(colour = "black")) +
  theme(legend.position = "none") +
  ylim(0, max(Ros.887$DataPoint)+2)
ggsave(filename = paste0("./Scripts/2023_dataset/Batch_3/Plots/SumNormInt.DESI_887_YvO.", Sys.Date(), ".pdf"), width = 3, height = 4.5, useDingbats = FALSE)
```
